My wonderful paper

### 2024-11-13 _manuscript/_tex/index.tex
Section 3.2 toy example'

The expectations can be thought of as a set of unit tests used to
validate the results of data analysis. By specifying a range of values
for these tests, multiple versions of the dataset can be generated to
satisfy different sets of plan expectations. This allows us to present
what we called the ``result universe'' -- the complete set of possible
results that can be obtained from one data analysis process. By
visualizing the result universe, data analysis consumers can observe how
changes in expectations affect the results and the range of alternative
outcomes that could arise under different conditions. This enables them
to evaluate the outcomes based on their own plan expectations and gain a
broader perspective on how the actual results produced by analysts fit
within this spectrum of possibilities, promoting transparency and trust
in analysis.

@@

Let's think about a 5-day step count. You make a resolution to walk on
average 5000 steps a day (your expectation) and using an app to record
your step count. After 5 days, the app tells you've walked on average
8000 steps.

It is easy to come up with reasons why an 8000 average step is resulted
based on common sense:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  you may run a 10k on day 1, resulting a high step count on the day
  (outlier on the right).
\item
  you left your phone at home on day 3, resulting a zero or minimal step
  count on the day (outlier on the left).
\item
  you may realise the step count may increase since you were in a hiking
  trip in the last five days (average shift).
\end{enumerate}

Based on these reasons, you may devise a set of unit tests to check the
step count data, i.e.~check the maximum and minimum step count, check
the difference between each day.

\begin{itemize}
\tightlist
\item
  If the daily count looks like c(4000, 5000, 5500, 5500, 20000), the
  maximum check will flag the data for investigate the maximum. The
  difference between days test will also flag the data
\item
  If the daily count looks like c(20000, 20000, 20000, 20000, 20000),
  the maximum check will flag the data for investigate the maximum.
\end{itemize}

Some part of the space is impossible: c(0, 4000, 5000, 5500, 5500) is
flagged by the minimal tests but won't cause an average of 8000 average
step.

The statistical procedure of averaging 5 numbers ``around 5000'' to get
a mean of 5000 is \emph{consistent} meaning if all the numbers are
around 5000, we are guaranteed to get a mean around 5000. We could
devise 5 unit tests to check each number. Since you're more familiar
with your daily life, you may realise the step count may increase since
you were in a hiking trip in the last two days. This may prompt you to
check the step count.

In a data analysis, it is not practical to check every entry of the
data, a similar strategy of devising tests to check for

\begin{itemize}
\tightlist
\item
  The combination of unit tests are not unique
\item
  The unit tests provide guidance for diagnosing the results, but are
  not red flags: c(2000, 2000, 5000, 8000, 8000) will likely to fail the
  max diff test but receive a within expectation mean.
\end{itemize}

>>>>

@@

Consider a 30-day step count goal. Suppose you resolve to walk at least
8,000 steps each day, using an app to record your daily step count. Your
target average is 9,000 steps per day, with some ``lows'' day, where you
walk around 4,000 steps and ``high'' days where you reach about 12,000
steps. After 30 days, you check how many times your step count fall
below 8,000 and aim for no more than five days under this threshold.

To simulation this data, three normal distributions with different means
are used for the daily step counts: \(\mathcal{N}(4000, 200)\) for low
days, \(\mathcal{N}(12000, 200)\) for high days, and
\(\mathcal{N}(9000, 300)\) for typical days. The number of low and high
days can be simulated from a Poisson distribution with \(\lambda = 4\).
Figure~\ref{fig-step-count} displays the number of days with fewer than
8,000 steps across 300 simulated 30-day periods.

In this scenario, the outcome expectation is that the number of days
with a step count below 8,000 will be no more than five. To diagnose
potential reasons why this outcome expectation might fail, we can
establish a few plan expectations. For example, if the average step
count is too low, this may suggest there are too many low days,
potentially lead to an unexpected outcome. Similarly, we can also check
the quantile of the step count, if more than a third of the days fall
below 8,000, this could indicate an excess of low-count days.
Additionally, we may may expect the standard deviation of the step count
not to be overly large.

These considerations yield the following three unit tests as plan
expectations:

\begin{itemize}
\tightlist
\item
  test1: the test fails if the mean step count is below 8,200
\item
  test2: the test fails if the 30th percentile of the step counts is
  below 8,200
\item
  test3: the test fails if the standard deviation of the step
  countsexceeds 2,500.
\end{itemize}

\phantomsection\label{cell-fig-step-count}
\begin{figure}[H]

\centering{

\includegraphics{index_files/figure-pdf/fig-step-count-1.pdf}

}

\caption{\label{fig-step-count}Number of days with fewer than 8,000
steps across 300 simulated 30-day periods. The orange bars indicate
instances where the count exeeds five days, representing an unexpected
outcome in this scenario.}

\end{figure}%

### 2024-11-13 index.qmd
Section 3.2 toy example'

The expectations can be thought of as a set of unit tests used to validate the results of data analysis. By specifying a range of values for these tests, multiple versions of the dataset can be generated to satisfy different sets of plan expectations. This allows us to present what we called the "result universe" -- the complete set of possible results that can be obtained from one data analysis process. By visualizing the result universe, data analysis consumers can observe how changes in expectations affect the results and the range of alternative outcomes that could arise under different conditions. This enables them to evaluate the outcomes based on their own plan expectations and gain a broader perspective on how the actual results produced by analysts fit within this spectrum of possibilities, promoting transparency and trust in analysis.

@@

Let's think about a 5-day step count. You make a resolution to walk on average 5000 steps a day (your expectation) and using an app to record your step count. After 5 days, the app tells you've walked on average 8000 steps.

It is easy to come up with reasons why an 8000 average step is resulted based on common sense:

1. you may run a 10k on day 1, resulting a high step count on the day (outlier on the right).
2. you left your phone at home on day 3, resulting a zero or minimal step count on the day (outlier on the left).
3. you may realise the step count may increase since you were in a hiking trip in the last five days (average shift).

Based on these reasons, you may devise a set of unit tests to check the step count data, i.e. check the maximum and minimum step count, check the difference between each day.

- If the daily count looks like c(4000, 5000, 5500, 5500, 20000), the maximum check will flag the data for investigate the maximum. The difference between days test will also flag the data
- If the daily count looks like c(20000, 20000, 20000, 20000, 20000), the maximum check will flag the data for investigate the maximum.

Some part of the space is impossible: c(0, 4000, 5000, 5500, 5500) is flagged by the minimal tests but won't cause an average of 8000 average step.

The statistical procedure of averaging 5 numbers "around 5000" to get a mean of 5000 is *consistent* meaning if all the numbers are around 5000, we are guaranteed to get a mean around 5000. We could devise 5 unit tests to check each number. Since you're more familiar with your daily life, you may realise the step count may increase since you were in a hiking trip in the last two days. This may prompt you to check the step count.

In a data analysis, it is not practical to check every entry of the data, a similar strategy of devising tests to check for

- The combination of unit tests are not unique
- The unit tests provide guidance for diagnosing the results, but are not red flags: c(2000, 2000, 5000, 8000, 8000) will likely to fail the max diff test but receive a within expectation mean.

>>>>

<!-- The expectations can be thought of as a set of unit tests used to validate the results of data analysis. By specifying a range of values for these tests, multiple versions of the dataset can be generated to satisfy different sets of plan expectations. This allows us to present what we called the "result universe" -- the complete set of possible results that can be obtained from one data analysis process. By visualizing the result universe, data analysis consumers can observe how changes in expectations affect the results and the range of alternative outcomes that could arise under different conditions. This enables them to evaluate the outcomes based on their own plan expectations and gain a broader perspective on how the actual results produced by analysts fit within this spectrum of possibilities, promoting transparency and trust in analysis.   -->

@@

Consider a 30-day step count goal. Suppose you resolve to walk at least 8,000 steps each day, using an app to record your daily step count. Your target average is 9,000 steps per day, with some "lows" day, where you walk around 4,000 steps and "high" days  where you reach about 12,000 steps. After 30 days, you check how many times your step count fall below 8,000 and aim for no more than five days under this threshold.

To simulation this data, three normal distributions with different means are used for the daily step counts: $\mathcal{N}(4000, 200)$ for low days, $\mathcal{N}(12000, 200)$ for high days, and $\mathcal{N}(9000, 300)$ for typical days. The number of low and high days can be simulated from a Poisson distribution with $\lambda = 4$. @fig-step-count displays the number of days with fewer than 8,000 steps across 300 simulated 30-day periods.

In this scenario, the outcome expectation is that the number of days with a step count below 8,000 will be no more than five. To diagnose potential reasons why this outcome expectation might fail, we can establish a few plan expectations. For example, if the average step count is too low, this may suggest there are too many low days, potentially lead to an unexpected outcome. Similarly, we can also check the quantile of the step count, if more than a third of the days fall below 8,000, this could indicate an excess of low-count days. Additionally, we may may expect the standard deviation of the step count not to be overly large.

These considerations yield the following three unit tests as plan expectations:

  - test1: the test fails if the mean step count is below 8,200
  - test2: the test fails if the 30th percentile of the step counts is below 8,200 
  - test3: the test fails if the standard deviation of the step countsexceeds 2,500.

```{r}
#| label: fig-step-count
#| fig-cap: Number of days with fewer than 8,000 steps across 300 simulated 30-day periods. The orange bars indicate instances where the count exeeds five days, representing an unexpected outcome in this scenario. 
set.seed(1234)
step_count <- tibble(id = 1:300) |> 
  rowwise() |> 
  mutate(
    small_n = rpois(1, 4),
    large_n = rpois(1, 4),
    norm_sd = rlogis(1, 800, scale = 100),
    norm_n = 30 - small_n - large_n,
    step = list(round(c(rnorm(small_n, 4000, 200), 
                        rnorm(large_n, 12000, 200), 
                        rnorm(norm_n, 9000, 300)
                        ), 0)),
    less_than_8000 = sum(step < 8000), 
    unexpect = ifelse(less_than_8000 > 5, 1, 0),
    test1 = ifelse(mean(step) < 8200, 1, 0),
    test2 = ifelse(quantile(step, 0.3, na.rm = TRUE) < 8300, 1, 0),
    test3 = ifelse(sd(step) > 2500, 1, 0),
    ) |>
  ungroup()

# step_count |> 
#   dplyr::select(id, step) |> 
#   unnest(step) |>
#   mutate(day = rep(1:30, 300)) |> 
#   ggplot(aes(x = step, y = id)) + 
#   geom_point(alpha = 0.5) + 
#   theme_bw()

step_count |> 
  ggplot(aes(x = less_than_8000, fill = as.factor(unexpect))) + 
  geom_histogram(binwidth = 0.5) + 
  scale_x_continuous(breaks = seq(0, 13, 1)) +
  scale_fill_brewer(palette = "Dark2") + 
  xlab("Number of days with step count less than 8000") + 
  theme_bw() + 
  theme(legend.position = "none")
```
